\SW: II. Did we beat the lens finding robots in CFHTLS?

### 2014-08-21 doc/sw-cfhtls.tex
Copied files from simpaper directory

\begin{abstract} 
We present X new strong gravitational lens candidates found in the CFHT
Legacy
survey by the \sw collaboration.

@@

{\it Describe various lens finding algorithms and the motivation for this
particular project with \cfhtls. Already searched by robots: enables comparison of techniques. Lenses not
yet found by robots, detectable by humans? }

In Paper I, we describe the \sw website, an online system that enables
crowd-sourced detection of gravitational lenses. In this paper, we
describe our first lens search with \sw in the \cfhtls data which we
refer to as Paper II.

@@

The \cfhtls data has been searched for lenses using semi-automated
algorithms, primarily, in the $g$ band which is optimal for finding
lenses. Here, we briefly mention the lens samples which were known to
the authors prior to the lens search with \sw.
The work
by \citet{Gavazzi2013} looks for bluish features around bright XXX
galaxies and has found XXX lens candidates at galaxy-scales.  The arc
finding lens search \citet{More2012} has found 127 lens candidates by
looking for arcs in $g$ band alone. This sample has both galaxy and
groups scale lens candidates. The \citet{Sygnet2009} is a search for
edge-on disk lenses done by selecting the galaxy's profile from the
Sextractor's output in the $i$~band and has 18 lens candidates.

For the purpose of transparency, the volunteers participating in
\sw-\cfhtls lens search were made aware of these known lens samples. Images
containing the lenses from these lens samples were labelled so in the
forum Talk where volunteers have the opportunity to discuss their
findings with other volunteers and the science team. XXX some references
to Paper I ?

@@

\section{Results}
\label{sec:results}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Recovery of lenses}
\label{sec:results:reco}

{\it Describe which lenses were recovered and explain any cases missed by
citizen scientists}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
%
%\begin{longtable*}{l c c c c c c c l}
%\caption{ \label{tab:all} 
%Characteristics of lens candidates from the \SA sample. }\\
%ID & RA & Dec &  $i$ & $z_{\rm phot}$ & $R_{\rm A}$ & Rank & Type & previous ID  \\
%    & (hms) & (dms) &  (mag) &   & (arcsec) &  &  & \\ [1.5ex] \hline \\ 
%\endfirsthead
%
%ID & RA & Dec &  $i$ & $z_{\rm phot}$ & $R_{\rm A}$ & Rank & Type & previous ID  \\
%    & (hms) & (dms) &  (mag) &   & (arcsec) &  &  & \\ [1.5ex] \hline \\ 
%\endhead
%
%\hline  
%\multicolumn{11}{p{15cm}}{
%{\bfseries
%{\bf m:} This galaxy falls within the masked region as per the catalog from
%which the magnitudes and the redshift are extracted.  {\bf z:} The magnitudes
%and/or redshift are not from the catalog instead are measured by the author.
%}  }\\
%
%\endlastfoot
%   SW1  & 02:01:21.89 & $-$09:15:15.09 & 19.90 &  0.46 &  2.2 &   2.0 & A &  SA18 \\ 
%   SW2  & 02:02:10.50 & $-$11:09:11.68 & 17.80 &  0.48 &  5.0 &   3.7 & A &  SA87 \\ 
%\end{longtable*}
%
\subsection{Stage 1 Classification}
\label{sec:results:stage1}

@@

\subsection{Discovery of lenses}
\label{sec:results:disc}

\subsection{\cfhtls completeness and purity}
\label{sec:results:sample}

@@

We looked at the locations of simulated and real lenses from our data
those were missed by \sw compared to the locations of the lenses that
got detected. We do not find any obvious dependency in the rate of
detections as a function of the position of a given lens for both
simulated and real lens sample. Thus, the completeness of the lens
sample is not affected by whether a lens is located close to the border
or well within the center (see  Figure~\ref{fig:comppos}).  {\bf TBD: redo the completeness vs posn
figure using the final real lens sample}

\begin{figure*}
\begin{center}
\includegraphics[scale=0.95]{sw-cfhtls-figs/completeness_pos.pdf}
\caption{ \label{fig:comppos}
Completeness as a function of location of lenses. Simulated (left) and
real lens (right) samples are shown. {\bf: TBD - Add a third color for
P$>$ Pthresh which will be defn of "accepted"} Irrespective of whether the
lenses are detected or rejected and whether they are simulated or real,
there is no obvious dependency on where the lenses are located. 
}
\end{center}
\end{figure*}

@@

\item We found a sample of new gravitational lens candidates. An expert-graded sample of size N, divided among several types.
The SW lenses are different from the robotically (RingFinder and ArcFinder) detected lenses, in the following ways.

% An expert-graded sample of size N, divided among several types.

>>>>

\begin{abstract}

The CFHT Legacy Survey has been searched for strong lenses with
semi-automated algorithms both at galaxy and groups scales. With the aim
of improving these lens finding robots, we carry out a blind lens search
in the complete \cfhtls WIDE survey with \sw. We describe the training
sample used for both training the citizen scientists that participated
in \sw and calibrating their performance. We generate realistic looking
simulated samples of lenses both at galaxy and group-scales as part of
this training sample. We present 74 new strong gravitational lens
candidates discovered from the \sw-\cfhtls search. Furthermore, we
compile a sample of false positives which can be used for robot testing,
first of its kind for the lensing community. The simulated sample is
also used to make a comparison between human performance against the
robots. We find that the XX per cent of the known lens sample is
recovered and the new lens sample has XX completeness with respect to
the simulated sample. We make available the following data products:
simulated sample, simulation code and the false positives sample at XXX.

@@

%{\it Describe various lens finding algorithms and the motivation for this particular project with \cfhtls. Already searched by robots: enables comparison of techniques. Lenses not yet found by robots, detectable by humans? }

The last few decades have seen a rise in the discoveries of strong
gravitational lenses owing to the plethora of interesting applications
they have in astrophysics and cosmology. Strong lenses are routinely
used to probe the dark matter distribution from galaxy (ref) to cluster
scales (ref), to  study distant young galaxies by using the lensing
magnification as a natural telescope (ref), to test the cosmological
model by constraining cosmological parameters such as the Hubble
constant (ref) and dark energy (ref) and many more. Even though strong
lenses are rare, since a foreground massive object needs to be
sufficiently aligned with a distant background source to produce
multiple images, systematic lens searches have led to discovery of over
500 lenses till date (XXX add mld url).

Rarity of lenses implies searching for them is a painstaking task.
Efficient automated methods are thus imperative to finding a reasonably
complete and pure sample of strong lenses.

Since the inception of the first citizen science project, Galaxy Zoo, to
classify galaxy morphologies \citep{LintottEtal2008}, several astronomy
and non-astronomy projects have been launched by the Zooniverse leading
to many interesting project. For example, PlanetHunters
has discovered XXX transiting exoplanets \citep{SchwambEtal2012},
Supernova XXX add ref and XXX add ref

In Paper I, we describe \sw, an online system that enables crowd-sourced
detection of gravitational lenses. In this paper (referred to as Paper
II), we describe our first lens search with \sw in the \cfhtls data.

@@

The \cfhtls data has been searched for lenses using semi-automated
algorithms, primarily, in the $g$ band which is optimal for finding
lenses. Here, we briefly mention the lens samples which were known to
the authors prior to the lens search with \sw.

At galaxy-scales, there are two main types of lens searches. The
RingFinder \citet{Gavazzi2014} was used for finding compact rings or arcs
around centers of isolated and massive early-type galaxies. By
subtracting the PSF-matched i-band images from the g-band, the
algorithm looks for excess flux in the bluer g-band which is typical of
the distant lensed star forming galaxies. An object detector measures
the properties of these residual blue features and candidates meeting a
length-width ratio and tangential alignment criteria are then visually
inspected to form the final sample. This rings sample consists of 39
confirmed lenses and 17 promising candidates.

The second galaxy-scale lens search was to find edge-on galaxy lenses in
the \cfhtls \citep{Sygnet2010} by selecting the galaxy's profile from the
output of {\sc Sextractor} in the $i$~band and with low inclination
angle. This sample has about 3 promising and a total of 18 lens
candidates.

On the other hand, the {\sc arcfinder} \citet{More2012} was used for
finding blue arc-like features in the complete \cfhtls data withouth any
pre-selection on the type of the lensing object. The search was carried
out in the g-band which is the most efficient wavelength to find typical
lensing galaxies. This sample, called the SARCS, has 55 promising and a
total of 127 lens candidates and consists of both galaxy and
groups/cluster scale lens candidates. Arc finding is better suited for
lensed images or arcs with larger image separations i.e. more massive
systems like groups and clusters. Thus, the SARCS sample has mostly
groups/cluster-scale lenses and a few galaxy-scale lenses.

For the purpose of transparency, the volunteers participating in
\sw-\cfhtls lens search were made aware of these known lens samples. Images
containing the lenses from these lens samples were labelled so in the
forum Talk where volunteers have the opportunity to discuss their
findings with other volunteers and the science team.

XXX refer Paper I ?

@@

\section{Results}
\label{sec:results}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Discovery of new lenses from \sw}
\label{sec:results:disc}

%
%\begin{longtable*}{l c c c c c c c l}
%\caption{ \label{tab:all} 
%Characteristics of lens candidates from the \SA sample. }\\
%ID & RA & Dec &  $i$ & $z_{\rm phot}$ & $R_{\rm A}$ & Rank & Type & previous ID  \\
%    & (hms) & (dms) &  (mag) &   & (arcsec) &  &  & \\ [1.5ex] \hline \\ 
%\endfirsthead
%
%ID & RA & Dec &  $i$ & $z_{\rm phot}$ & $R_{\rm A}$ & Rank & Type & previous ID  \\
%    & (hms) & (dms) &  (mag) &   & (arcsec) &  &  & \\ [1.5ex] \hline \\ 
%\endhead
%
%\hline  
%\multicolumn{11}{p{15cm}}{
%{\bfseries
%{\bf m:} This galaxy falls within the masked region as per the catalog from
%which the magnitudes and the redshift are extracted.  {\bf z:} The magnitudes
%and/or redshift are not from the catalog instead are measured by the author.
%}  }\\
%
%\endlastfoot
%   SW1  & 02:01:21.89 & $-$09:15:15.09 & 19.90 &  0.46 &  2.2 &   2.0 & A &  SA18 \\ 
%   SW2  & 02:02:10.50 & $-$11:09:11.68 & 17.80 &  0.48 &  5.0 &   3.7 & A &  SA87 \\ 
%\end{longtable*}
%

\begin{figure*}
\begin{center}
\includegraphics[scale=1.5]{sw-cfhtls-figs/lenscand_2p0_3p0.pdf}
\caption{ \label{fig:lch1}
Sample of promising lens candidates ($R_{avg}>=2.0$)
}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[scale=1.5]{sw-cfhtls-figs/lenscand_2p0_3p0_1.pdf}
\caption{ \label{fig:lch2}
contd. - Sample of promising lens candidates ($R_{avg}>=2.0$)
}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[scale=1.5]{sw-cfhtls-figs/lenscand_1p3_1p7.pdf}
\caption{ \label{fig:lcl}
Sample of low probability lens candidates ($1.0<R_{avg}<2.0$)
}
\end{center}
\end{figure*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Recovery of known \cfhtls lenses from \sw}
\label{sec:results:reco}

{\it Describe which lenses were recovered and explain any cases missed by
citizen scientists}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{\cfhtls completeness and purity}
\label{sec:results:sample}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\subsection{Stage 1 Classification}
\label{sec:results:stage1}

@@

@@

{\bf TBD: 
add section on offline analysis - extra lenses found etc;  
expand FP section;  
mention EGS moustakas and other lens samples; 
comment on any lensed quasars, on any exotic lenses;
check if some candidates were detected because they were hidden
underneath the sims ie. from the D11;

}

We looked at the locations of simulated and real lenses from our data
those were missed by \sw compared to the locations of the lenses that
got detected. The real lens sample consists of a total of 383 candidates
which have $P_l>2.e-5$ and received a rank of $R_{avg}>0$ from the
expert. We do not find any obvious dependency in the rate of detections
as a function of the position of a given lens for both simulated and
real lens sample. Thus, the completeness of the lens sample is not
affected by whether a lens is located close to the border or well within
the center (see  Figure~\ref{fig:comppos}).

\begin{figure*}
\begin{center}
\includegraphics[scale=0.95]{sw-cfhtls-figs/completeness_pos.pdf}
\caption{ \label{fig:comppos}
Completeness as a function of location of lenses. Simulated lenses (left) and
real lens candidate (right) are shown. {\bf: TBD - Add a third color for
P$>$ Pthresh which will be defn of "accepted"} Irrespective of whether the
lenses are detected or rejected and whether they are simulated or real,
there is no obvious dependency on where the lenses are located. 
}
\end{center}
\end{figure*}

@@

\item We found a sample of new gravitational lens candidates. An
expert-graded sample of 74 with 44 promising and 30 low probability
candiates.

The SW lenses are different from the robotically (RingFinder and ArcFinder) detected lenses, in the following ways.
XXXXX
